---
title: "Verifying FMF model outputs"
output: word_document
---

```{r, echo=FALSE}
# setwd("../../") # only run this in 'knit' mode - sets wd correctly
```


# Loading the data

The 1 Gb file was first loaded. To make this manageable, it was shortened to
100,000 individuals. This is OK because the list contains the entirety of
the population in one zone, then the next, in order.

```{r, eval=FALSE}
ids <- read.csv("data-big/final-pop-stephen/RL1_population.csv")
head(ids)
ids <- ids[1:50000,]
ids$X.PersonID <- as.factor(ids$X.PersonID)

# load model matrix individual data
ind_catid <- cbind(X.PersonID = as.character(ind$PersonID), ind_cat)
ind_catid$X.PersonID <- as.factor(ind_catid$X.PersonID)

save(ind_catid, ids, cons, file = "~/repos/IPF-performance-testing/models/experimental/test-fmf.RData")
```

# Preparing the individual-level data

The next stage is to add 'id' classes to the pre-saved individual-level data.
Checks were undertaken to ensure that the same IDs were present in the
individual-level data loaded previously.

```{r}
load("~/repos/IPF-performance-testing/models/experimental/test-fmf.RData")
head(ind_catid[1:5])

library(dplyr)
ids_all <- inner_join(ids, ind_catid)
summary(ids_all) # test to see the join has worked 
```

The `join` function above added all the attribute data for the individuals
(in 'model matrix' form, consisting of a column for each category, for ease
of counting). This allows the fit between the constraints and the aggregated
versions of the spatial microdata to be tested. We don't aggregate the
entire dataset, we aggregate by zone.

# Grouping the microdataset by zone

```{r}
ags <- aggregate(ids_all[-(1:2)], by = list(ids_all$ZoneID), FUN = sum)
head(ags)
head(rowSums(ags[-1]), 10)
head(rowSums(cons), 10)
ag <- ags[1:6, -1]

cor(as.numeric(as.matrix(ag)), as.numeric(as.matrix(cons[1:6, ])), )
colSums(ag[,1:6]) # Number of divorced, married, single and widowed from FMF
colSums(ag[,1:6]) - colSums(cons[1:6,1:6])
```

The above demonstrates the perfect fit attained between
the aggregated spatial microdata and the constraint tables.

